### 2020-01-15 convivial-2020/tex/manuscript.tex
Abstract criteria

\documentclass[english,submission]{programming}
\usepackage[backend=biber]{biblatex} % Use Biblatex
\addbibresource{example.bib}

\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete, do, else,
  finally, for, function, if, in, instanceof, new, return, switch, this, throw,
  try, typeof, var, void, while, with, let},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  sensitive=true
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\paperdetails{
  perspective=engineering,
  area={General-purpose programming}
}

\title{What does it take to create with domain-appropriate tools?}
\subtitle{A case study on the ``OROM'' system}
\author{Joel Jakubovic}
\affiliation{University of Kent, Canterbury}

\keywords{paper, showcase, lorem ipsum}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002944.10011122.10003459</concept_id>
<concept_desc>General and reference~Computing standards, RFCs and guidelines</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{General and reference~Computing standards, RFCs and guidelines}`

\maketitle

\begin{abstract}
  This paper shows...
\end{abstract}

  \hypertarget{introduction}{%
  \section{Introduction}\label{introduction}}

  As someone who can code, I have already passed the first and most
  important hurdle for making full use of the potential of my computer.
  However, even in this supposedly empowered state, I am still far away
  from feeling the relationship between myself and software as between
  artisan and material, free to shape it into any form with effort
  proportional to complexity.

  One would have thought that software-creation acts like hypothetical
  super-intelligent AI. That is: even though we start from a primitive
  base in the 50s (or even today), there would surely be a recursive
  process of self-improvement, building better software-creation tools
  with the existing ones, until an ``expressivity singularity'' where
  software becomes a workable material as described.

  However, this didn't happen. Or at least, it is happening glacially
  slowly. The brute fact is that whenever you want to create software, you
  go to a text editor and figure out how to translate your design into
  that. The text editors, being software, were written with the help of
  previous text editors, and so on. It's undeniable that text editors have
  improved, even if you think it peaked with Emacs. We just don't seem
  able to go beyond them where it matters, such as visual domains
  ill-fitted to monospaced ASCII.

  Amdahl's Law generalises the following idea: even when you spend hours
  of effort doubling the performance of a component used 1\% of the time,
  your reward is a system overall improved by a mere 0.5\%. Now, text
  coding is certainly ubiquitous, the 99\% case in programming, so you
  might wonder where I'm going with this. Well, a small improvement to
  text editing, if adopted by everyone, certainly does have a massive
  \emph{intermediate} effect --- but this only \emph{matters} to the
  extent that text was helping the programmers in the first place. If my
  goal is to draw or animate pictures, or create a digital synth from a
  frequency spectrogram, then giving me the ability to auto-indent my SVG
  markup is rather underwhelming as a productivity increase, as it doesn't
  target the core of the enterprise that makes it so hard.

  My experience of coding, most of the projects requiring shapes (such as
  GUIs), leads me to conclude that no matter how much I improve my skill
  at a particular language, knowledge of libraries or even general coding
  ability, my predicament stays the same. Our basic method of creating
  software is optimised for an ever-diminishing proportion of the software
  we actually want to make; ill-optimised for the graphics, layout,
  interactivity and and basic physics --- more on this later --- that we
  usually require.

  Whenever I work on these I feel stuck in a box I know I can never escape
  from: that box is the text editor, a fixed conduit through which all
  \emph{fundamental} changes to my program must pass. It's not a part of
  the system I am building, so I can't even make use of features of the
  thing I'm developing, to make its own development easier.

  Surely the trick is to \emph{use} coding to build something \emph{better
  than it}. And then use that, to build something even better. But there
  is an enormous breadth and depth of philosophies here, along with many
  failed historical attempts to do better --- or at least, ones that
  failed to catch on. But even worse than this is that in my very
  \emph{language} here I am making the same mistake as the text editor ---
  speaking in unqualified terms of ``better'' and ``worse'' as if there
  really is a one-size-fits-all solution to software creation!

  Of course what we \emph{really} want is the ability for people to create
  \emph{in the way that they think is best} in their particular context
  --- to equip them to feasibly create the tools that suit them for the
  thing they want to make. And second-order tools that suit them for
  making the first-order tools, etc. It would do no good to replace
  text-imperialism with anything-else-imperialism, which is one
  interpretation of calls for alternatives. This dream goes beyond
  familiar senses of what constitutes a ``craft'', as far as a strong
  melding of tool and material.

  Certainly in the real world there is only one ``material'' --- matter
  --- and humanity as a whole does create ``higher-order'' tools. A
  (future) software craft, however, could be expected to give this power
  to \emph{individuals} instead of the community alone. The
  one-size-fits-all approach is an understandable consequence of few
  designers catering to many users --- perhaps due to the necessity to
  specialise in topics, languages, libraries, and tools. Their complexity
  is part of the general difficulty of software creation, exacerbated by
  the one-size-fits-all approach, closing the circle.

  What we need is some system that not only lets us create software in a
  way that is ``close to the problem domain'' as decided by the
  user-developer, but also can augment or change itself to adapt to a
  different ``way of creating''. Existing systems seem to only have one of
  these properties without the other. Smalltalk and LISP try to minimise
  arbitrary commitments of language \emph{semantics} to be adaptable, but
  their being textual languages is a fairly tough commitment to break out
  of. And it is not so hard to make a specific, \emph{hard-baked} visual
  or alternative programming tool --- but it is hard to make it
  re-programmable \emph{without} having to go back to \emph{its} textual
  source code.

  Again, if someone wants to type out pictures in ASCII, let them ---
  whether they do it for a challenge, or even if they find that more
  natural for themselves. But equally, if I want to do it another way,
  then please give me that affordance. This is how I segue into the
  software artefact for which I have been attempting to build a natural
  representation. True, it is a programmer's artefact, but it is still
  representative of what any normal person has to do, insofar as:

  \begin{enumerate}
  \def\labelenumi{\alph{enumi})}
  \tightlist
  \item
    Wanting to create a piece of software (for whatever reason)
  \item
    Having in mind a natural way to represent it as it's being built.
  \end{enumerate}

  \hypertarget{the-orom-system}{%
  \section{The OROM system}\label{the-orom-system}}

  When I first read the paper Open, Reusable Object Models, I was hooked
  on its idea of a small but expressive starting system that could be
  self-improved into anything. It describes a late-bound\footnote{Fewer
    commitments; more things determined by runtime conditions},
  Smalltalk-style\footnote{OOP with more emphasis on object instances and
    messaging (methods) as in a distributed system; less emphasis on class
    hierarchies implementing traditional data structures} objects and
  messaging environment that the authors call ``Id'' --- but which I refer
  to as OROM, for pronunciation and Googleability.

  In summary, an object is a key-value mappings of state which can change
  as a result of messages received by it. A message is sent by first
  \emph{bind}-ing its name to its implementation, i.e.~specific code which
  is then run in the context of the receiver. ``Bind'' is accomplished via
  another message send, this time to the receiver's \emph{vtable}, or
  ``behaviour''. This is another object whose state defines a particular
  way of associating ``message name'' to ``implementation code''. This
  continues up the vtable-chain, and terminates at a base case. The system
  as a whole is ``bootstrapped'' into existence by its initialisation
  code. Each step of this code makes use of any parts of the system set up
  by the previous steps.

  The paper itself consists of mostly prose, several code listings, and
  full C sources for a sample implementation at the end. It also provides
  several diagrams. But to understand it, I repeatedly found myself
  drawing \emph{extra} diagrams. For example, the first acts of the
  running system boil down to initialising the three or so objects. This
  consists of allocating memory, interpreting it as a C struct and then
  filling in fields in a mundane manner. I had great difficulty following
  the specifics in my head, but when I drew tables in the style of their
  diagrams I readily saw what what going on. I personally would have
  preferred these to have been in the paper in the first place, but I am
  not necessarily representative of those who read it. The
  ``one-size-fits-all'' approach is perhaps unavoidable for static media
  like print.

  Other areas like messaging semantics could still be quite confusing.
  After all, should we expect to be able to predict the entire future
  evolution of a dynamical system from a static description of its initial
  state, such as source code? Is this not why debuggers exist? Having
  ``de-compiled'' English text and C source code into object diagrams, it
  is a shame to have to compile it all back to struct member assignments.

  Worst of all, the reference system does not even have text I/O when it
  is run, let alone some sort of GUI. That is, the (un)intended user
  interface for this project is a C debugger! Faced with the necessity of
  adding \emph{some} UI, it seemed a waste of effort to end up with a
  system that must be continually polled for its current state at a
  terminal prompt. If I naturally think of this system as 2D tables, why
  can't that be how the running system looks? I do not have to keep
  polling my eyes for what state my diagrams are in.

  But further than that --- why can't the system be \emph{built} out of
  tables in the first place? Shouldn't this be the main takeaway from the
  amount of time we spend prototyping, explaining and designing software
  as diagrams on paper? Why must the ``natural representation'' be
  restricted to the finished product?

  Thus was my natural representation decided. My first attempt to make it
  a reality was a partial success: a webpage made of HTML tables, evolved
  via JavaScript.

  \hypertarget{orom-as-html-tables}{%
  \subsection{OROM as HTML tables}\label{orom-as-html-tables}}

  (OROM/HTML screenshot)

  I was grateful for the browser's management of graphical layout,
  resizable text fields, and keeping the DOM tree synchronised with what
  you actually see. This last property enabled me to make the decision to
  \emph{directly} encode much of the system state in the DOM, achieving
  basic liveness (the thing on the screen is the actual thing) for the
  keys and values of obj-dicts. For each obj-dict I used a two-column
  \texttt{\textless{}table\textgreater{}} within a
  \texttt{\textless{}div\textgreater{}}, matching the diagrams, and gave
  the rows CSS class names matching the keys for easy lookup. Sending
  messages relied on the JavaScript console, but existing values in text
  boxes (such as method implementations) could be edited directly. Also
  thanks to the browser, I could easily make text boxes and
  \texttt{\textless{}div\textgreater{}}s resizable as necessary.

  However, the choice of ordinary HTML as a substrate proved rather
  two-edged. The browser requires many features to do its job of rendering
  complex web pages. And sadly, as its client, I could only make use of
  those capabilities which the W3C had decided, at the time of authorship,
  were worth the effort exposing in JavaScript. For anything else, the
  browser is a black box, and this was very frustrating in the following
  case.

  \hypertarget{the-radical-concept-of-arrows-that-stay-on-the-shapes}{%
  \subsubsection{The Radical Concept of Arrows That Stay On The
  Shapes}\label{the-radical-concept-of-arrows-that-stay-on-the-shapes}}

  A key aspect of the OROM system is that there is an object \emph{graph}.
  That is, obj-dicts can have entries pointing to other obj-dicts, without
  restriction to a tree structure. Drawing arrows between things to denote
  this is a no-brainer, and I wanted it in my substrate for OROM to match
  the diagrams.

  A small problem to surmount first: even though I could hijack the
  \texttt{\textless{}table\textgreater{}} for its display properties, what
  element could I hijack to make arrows between arbitrary points? Luckily,
  there was SVG at my disposal, which could be persuaded to display over
  the HTML. However, another key feature of my intended substrate was to
  be able to rearrange and resize the boxes as desired. So I would also
  need to detect changes to the position and size of an element.

  Bizarrely, there is no such facility provided for HTML elements. This,
  despite the fact the browser \emph{needs} this functionality hence it
  must reside \emph{somewhere} inside the black box.

  Reluctantly, I stuck with my plan B: each object has a numerical ID and
  pointers are just fields containing a number, followed using a
  \texttt{deref()} function.

  At this point it was starting to look like a mistake not to have done
  the whole thing in SVG. I would gain full graphical freedom, though also
  lose some benefits of the browser's managing it on my behalf. I already
  knew from experience the surprising complexity of a DIY approach to
  layout, model-view updates, and interactivity. But such an exercise
  would be an opportunity to carefully observe this tedium, and
  crystallise some of my intuitions about why it is so consistently
  frustrating.

  \hypertarget{orom-as-files-and-directories}{%
  \subsection{OROM as files and
  directories}\label{orom-as-files-and-directories}}

  \begin{quote}
  I would probably put this after the SVG part or perhaps even later once
  you discuss more general principles (like the idea of polyfilling to
  make a domain substrate into a programmable system). Then you can give
  this as another example.
  \end{quote}

  \hypertarget{orom-as-svg-trees}{%
  \subsection{OROM as SVG trees}\label{orom-as-svg-trees}}

  (Screenshot of OROM/SVG showing boxes, code and arrows)

  In this version, obj-dicts are encoded as nested SVG
  \texttt{\textless{}rect\textgreater{}}s and other elements, reminiscent
  of diSessa's Boxer. This was a significant departure from the table
  representation, and even though SVG supports (some) nested HTML via
  \texttt{\textless{}foreignObject\textgreater{}}, I actually preferred
  the possibility of multiple levels of nesting.

  OROM/SVG more or less realises my desired substrate for implementing
  OROM. Needless to say, this version was far more challenging and took
  much longer to reach a satisfactory state. However, it is precisely this
  drudgery that brings me to a better understanding of this paper's
  question: \emph{what has it taken?} I shall discuss this in the form of
  broad patterns or themes that stand out to me from my development
  experience, hammered home by these OROM projects.

  \begin{quote}
  Why is OROM/SVG more challenging? I think it would be useful to sketch
  here, because it reveals some useful principles. There is more
  polyfilling (arrows and resizing is not in the SVG substrate itself),
  but at least it's possible to add those (unlike in HTML)
  \end{quote}

  \hypertarget{inevitable-requirements-of-most-software-and-the-work-we-must-do-to-meet-them}{%
  \section{Inevitable Requirements Of Most Software (and the work we
  must do to meet
  them)}\label{inevitable-requirements-of-most-software-and-the-work-we-must-do-to-meet-them}}

  As programmers, we have a maxim along the lines of: when you find
  yourself repeating the same thing over and over again, factor it into a
  ``thing'' and let the computer do the duplication. Some consider the
  ``design pattern'' to be simply what a ``language feature'' looks like
  when it's not part of the syntax.

  \begin{quote}
  The jump from very general abstraction idea to design patterns and
  language features in the above is not exactly clear - it would be good
  to help the reader understand why are you talking about this here.
  \end{quote}

  As I developed OROM/SVG I found myself implementing many such patterns.
  It seems that languages are often at the wrong level of abstraction for
  the requirements of modern software, necessitating the same boilerplate
  per project just to get up to basic functionality. This burden either
  falls on the author, or on the wider community to build and maintain
  higher-level frameworks, syntax extensions, etc.

  By contrast, these languages seem much better adapted to e.g.~batch-mode
  printf()-centric applications. That they fail for the \emph{common case}
  should be concerning. I will now present what seem to be the inevitable
  demands of this common case.

  \begin{quote}
  I'm not sure I understand what the above opening is trying to say - I
  suppose it might be because it is quite high-level/general. What is this
  section about more specifically? Is it about what certain kind of
  software consists of and what we need to build that? I would also be a
  bit careful about things like ``most modern software'' below. It would
  be good to instead characterize what kind of software are you interested
  in - end-user graphical systems? (a microservice or something like that
  does not need SVG\ldots{})
  \end{quote}

  \hypertarget{as-a-mere-consumer}{%
  \subsection{As A ``Mere Consumer''}\label{as-a-mere-consumer}}

  I begin by describing expectations of software that even ``end-users''
  hold, consigned as they are to more or less accept what we give them.

  \hypertarget{retained-mode-vector-graphics}{%
  \subsubsection{Retained-Mode Vector
  Graphics}\label{retained-mode-vector-graphics}}

  Most software is designed for the subset of people who have a colour
  display they can perceive. So right away it is going to require ways to
  draw coloured shapes. There are usually libraries for this (though note:
  not part of the language), but some only provide \emph{immediate mode}:
  commands to instantaneously rasterise pixels to a buffer. This is not
  enough for modern software, as we often expect animation, or at least to
  see things change as we interact. Most often we wish to see \emph{small
  changes} to the \emph{same} shapes, rather than completely different
  shapes altogether; the reification that this requires to persist between
  frames, is known as \emph{retained mode.}

  On this requirement, SVG fits the bill very well. Although it is not
  part of the JavaScript language \emph{per se}, it is a standard and
  widely supported technology of the Web \emph{platform}. We can observe
  that anyone with a browser \emph{in principle} has access to a powerful
  vector graphics editor --- just one with no GUI. I will return to this
  later.

  The SVG tree keeps the nice properties of the DOM, such as updating the
  display when shape parameters are changed. This is well-adapted to
  ``I/O-bound'' software like mine, where things change only in response
  to user input. If I wanted animation, this boils down to a regular
  ``advance simulation'' signal, and would require setting up some
  rendering loop. Alternatively, there is the W3C's chosen ontology of CSS
  animations, but see section ??.

  \hypertarget{basic-assumptions-about-physical-objects}{%
  \subsubsection{Basic Assumptions About Physical
  Objects}\label{basic-assumptions-about-physical-objects}}

  In any software making use of vector graphics, there is usually some
  level of ``physics'' expected by users. This need not be nearly as
  exhaustive as the word ``physics'' might imply, as in e.g.~physics
  engines for games; I feel it is important to recognise it for what it is
  instead of conceiving physics as an inherently complex thing to be found
  only in specialised simulations.

  \begin{quote}
  People might not find this easy to believe - but I think you can
  convince them by saying that pretty much all mobile phone apps have some
  sort of ``phone touch physics'' (with sliding, opening menus etc.).
  \end{quote}

  All humans learn a basic set of expectations about the things they see
  around them. Some of these, such as ``things fall down'', are not
  generally appropriate to software UIs --- perhaps because the screen has
  a role in our lives a more of a table work surface rather than a
  vertical wall, even if it is vertical in real life.

  The level of physics in software tends to not involve force, or mass, or
  very much at all, merely position and space. One thing that all usable
  software must do, for example, is not crush many visually complex
  shapes, such as lines of text, into the same region, since it becomes
  unreadable. Such concepts of ``solid objects do not intersect'' or
  ``only things at different layers may overlap'' are basic rules
  inherited from the real world of graphical presentation.

  I feel the need to point this out, because by default the computer does
  not know even the most obvious things about how space works, so we must
  laboriously algorithmise this intuitive concept. This is not only true
  in the case of 2-dimensional visual domains, but even in the
  1-dimensional case of memory allocation. The physics of 1D memory are
  something like this:

  \begin{itemize}
  \tightlist
  \item
    This number range 0000-FFFF is like a space, numbers = points
  \item
    Every point has at most one owner block
  \item
    These blocks are contiguous, finite ranges (i.e.~1D boxes)
  \end{itemize}

  Hence the boilerplate involved to realise this in any domain with
  something resembling space. Far from being a niche topic in games and
  graphics, spatial partitioning algorithms and data structures have
  surprising relevance to more ordinary software. Both memory allocation
  and graphical layout are essential to today's; shame that only one of
  those has been recognised as such and entered the runtime of modern
  languages.

  \hypertarget{translationally-rigid-bodies}{%
  \subparagraph{Translationally Rigid
  Bodies}\label{translationally-rigid-bodies}}

  When you have both a screen and a pointing device (e.g.~touch or mouse),
  immediately it becomes worth having ways to move things around in at
  least a minimally realistic way. We can debate the appropriateness of
  Direct Manipulation (DM) for various situations. But it does make a lot
  of sense in simple cases, such moving around subdivisions of space
  (e.g.~windows) or elements of a graphical design.

  In OROM, the obvious candidate for this is the obj-dicts, plus all
  nested boxes in OROM/SVG. If I move the top-level rect, then I expect
  its children to move with it. This is simply the translational physics
  of rigid bodies: this set of points all move together. Of course, proper
  rigid bodies might also rotate and have mass, but this is usually
  undesirable for UI elements.

  Translational rigidity can be expressed as the points X and Y always
  having the same displacement from each other. Or, when one point is
  moved, the rest also move by the same delta. This is a problem of
  preserving the relationship over time, which was a significant area of
  OROM/SVG.

  \hypertarget{maintaining-relationships-over-time}{%
  \subsubsection{Maintaining Relationships Over
  Time}\label{maintaining-relationships-over-time}}

  The model of state-mutation present in most imperative languages is what
  I call ``dumb'' state. The language provides an affordance to change any
  part of the state to a new value, but nothing else.

  What more could there be? Well, in \emph{every} software system there
  are certain rules, or ``invariants'' of internal consistency, such as
  ``translational rigidity'' above. Often, changes to any part of the
  system are permissible, but only if connected or dependent parts of the
  state change in response.

  The job of keeping track of who depends on whom can fall either on the
  programmer, or the computer. If the programmer has to do this, they can
  only go so far managing and simulating in their head. As systems grow
  more complex, it is only natural to try and make the computer more
  intelligent to do this work. What I am building up to is that whenever
  we (or I) consider constraints, or ``reactive'' programming, or the
  ``Observer'' pattern as things you only wheel out on special occasions,
  we only deceive ourselves into doing the same work less explicitly. It
  seems that such ``live state'' should be the expected common case for
  software development.

  If a platform does not provide a means to causally link and unlink bits
  of live state, then this must form part of the standard boilerplate.

  Such was the case in OROM/SVG. I implemented the system in an OOP
  fashion, and the Observable class is the most widely used. It wraps a
  current value and a list of subscribers, notifying them when it changes.
  It can be seen that this supplanted JavaScript's own ``dumb'' state by
  the frequency of the change() function throughout the code.

  Getting an object to follow the mouse pointer, for example when
  dragging, is conceptually very simple: an ``always equal'' relation. In
  OROM/SVG founded on live-state, this can be expressed in much the same
  way:

  \begin{verbatim}
  subscribe(object.position, pointer.position);
  \end{verbatim}

  (By default, an Observable A responds to a change from Observable B by
  adopting B's new value.)

  \hypertarget{nut-cracking-with-sledgehammers}{%
  \subsubsection{Nut-Cracking With
  Sledgehammers}\label{nut-cracking-with-sledgehammers}}

  Speaking about vector graphics, physics, layout and constraint
  maintenance might give the impression of high \emph{conceptual}
  complexity at the heart of even simple software. This is not quite true,
  which makes it all the worse that there is yet still immense
  \emph{incidental} or implementation complexity.

  We are conditioned to only think of these in their most general forms.
  The vector graphics I use in OROM/SVG are just rects, lines, circles and
  text; a fraction of the full capability of SVG. The ``basic physics'' I
  use is dwarfed by fully general 2D or 3D physics engines. The only
  layout algorithm I had the patience to implement was a simple way to
  expand a list of boxes to fit in a new child at the bottom; the
  affordance to place and size boxes \emph{manually} staves off the rest
  for the time being! Yet search for material on layout algorithms and it
  can seem like Fully General Linear Inequality Solvers like Cassowary are
  all there is.

  The inevitable requirements I suggest here, do \emph{not} necessitate
  Fully General anything. In fact, such representations would make it more
  cumbersome to express what I wanted in OROM. As the old wisdom goes,
  there's no point expressing a simple regex search as an arbitrary Turing
  machine; my boxes don't \emph{have} a moment of inertia or a mass and I
  don't \emph{need} a linear optimisation solver for my space management
  --- for the time being.

  Unfortunately, there doesn't seem to be much interest on the
  smaller-scale instances of these problems. I suggest the name
  ``geometric physics'' for the subset of physics I have mentioned.

  \begin{quote}
  I think the above section talks about 3 things - (i) it describes some
  of the things that you had to implement for OROM/SVG, (ii) it talks
  about some general patterns that arise there (like shapes and physics)
  and (iii) it mentions some implementation details. I found it a bit
  confusing to have these all mixed together. It reads a bit as if you
  were jumping from one random topic that you encountered when
  implementing this - but that makes the section very specific to your
  implementation. How can we make it clear that there are more general
  principles here? I think it might be better if you started with some
  larger example (hypothetical is fine!) to illustrate what kind of things
  you'd like to build - e.g. imagine you're building PowerPoint in OROM -
  what would be some of the things you'd need? (for both the end
  application and the system that you're using to build it). This would
  then nicely motivate the more specific topics (graphics, physics,
  relationships).
  \end{quote}

  \hypertarget{as-a-developer}{%
  \subsection{As A ``Developer''}\label{as-a-developer}}

  I now turn to some things that were perhaps inevitable, in any case
  useful, as the \emph{implementor} of OROM/SVG.

  \hypertarget{practicality-vs.-high-aesthetics}{%
  \subsubsection{Practicality vs.~High
  Aesthetics}\label{practicality-vs.-high-aesthetics}}

  There is the view that insists on provably-correct code written with
  strange Unicode symbols, and at the opposite end there is the attitude
  of ``get the job done'', high style be damned. Each has its merits, but
  for my task I quickly saw the latter route to be most suitable.

  My favourite example of this has been the ``encoding'' of obj-dicts as
  DOM trees containing particular arrangements of children. In OROM/HTML,
  I would gleefully get/set state by calling code like this:

  \begin{verbatim}
  Entity.prototype.addState = function(key) {
    let tr = document.createElement('tr');
    tr.setAttribute('class', key);

    let td = document.createElement('td');
    td.setAttribute('class', 'key');
    td.textContent = key;
    tr.appendChild(td);
    ...
  \end{verbatim}

  and in OROM/SVG the story is similar. It takes advantage of the powerful
  developer tools included in browsers (e.g.~the DOM inspector), and
  avoids the need to manually synchronise separate Model and View for such
  encoded state. With a little work, it also allowed the system a large
  degree of externalisability --- see Section ??. I consider this as
  practical, expedient, ``hijacking'' of these substrates because it
  either goes against their intended use (HTML tables for presentation
  rather than semantics) or are simply ``odd'' (using certain arrangements
  of shapes to ``encode'' machine-readable semantics).

  Contrast this with an approach that had to re-synthesise bits of the DOM
  whenever something changed. Admittedly this does still happen to a
  certain extent, but at least it was minimised where obvious.

  Another example of practicality was in the issue of positioning and
  sizing. In Section ?? I mentioned that I was able to sidestep some
  layout work by offloading it to the user. While building the system I
  tended to want to draw boxes into existence, and for that my brain's
  existing aesthetic algorithms can decide where they go. There are areas
  where the machine must be able to figure this out itself --- I don't
  want to be constantly bothered by prompts to place boxes as they are
  allocated by some running code --- but for one-offs it is a very
  convenient way to do a little less yak-shearing.

  \begin{quote}
  I think you could adapt the above discussion a bit and focus less on
  what you don't want (high modernism) and more on what you want - what is
  that? I think it's simplifying the problem along a different axis - to a
  certain domain with relatively simple interactions and aesthetics. (So
  rather than trying to make minimal formal calculus that's provably
  correct, you're trying to make minimal working interaction model that's
  interesting)
  \end{quote}

  \hypertarget{positioning-and-sizing}{%
  \subsubsection{Positioning and Sizing}\label{positioning-and-sizing}}

  The simple desire to move and resize boxes with the mouse motivated a
  lot of the live-state and geometric physics ideas. This problem could be
  considered a microcosm of OROM/SVG: what's a natural way I
  \emph{conceive} of this behaviour, and could I implement it that way?

  \begin{quote}
  The above point (moving motivated live state and physics ideas) would be
  a lot more obvious if you followed the structure with motivating example
  first followed by an analysis\ldots{}
  \end{quote}

  It starts with a consideration of translational rigidity. In its most
  primitive form, this is a relation between two points. Thus it is
  natural to draw a line or ``rod'' between them. It seems that this rod
  transmits changes in one of its endpoints directly to the other
  endpoint. Since they both feel the same deltas, the displacement vector
  between them is preserved.

  (Diagram of delta transmission)

  I like to see what I'm doing, so I wanted these rods to be visible and
  thus somehow \emph{present in the SVG}. This was not hard; in SVG I draw
  a \texttt{\textless{}line\textgreater{}}, though the background
  machinery of the Rod class does all the work. There is also a Point
  class that is used wherever manipulable points (SVG circles) are
  intended.

  Resizing (of boxes at least) could be achieved through rods that stay
  horizontal or vertical. In the language of ``small differences'' spoken
  by my live-state infrastructure, this is expressed as a rod
  ``transmitting deltas'' in the vertical and horizontal, ``absorbing''
  the other component into itself. Mirroring a DOM rect to these rods is
  as simple as subscribing its width and height to horizontal and vertical
  rods' length Observables. This way, boxes can be resized from whatever
  corner is convenient.

  (Box demonstrating all three rod types)

  I colour such ``half-rigid'' rods yellow and fully rigid rods red. The
  ``absorb all endpoint changes'' case is coloured green.

  \begin{quote}
  I don't think you explained what `half-rigid' means yet.
  \end{quote}

  Unfortunately, with these rods came possibly the most frustrating
  technical challenges of the entire system. Initially I hoped to move
  boxes as rigid bodies by temporarily making their border rods rigid.
  However, the four border rods form a cyclic graph, as rods are not
  directed. This, coupled with the unintended depth-first semantics of
  Observable notification (a result of JavaScript function calls in a
  loop), led to duplicate deltas applied twice and other nightmares. This
  is the tip of an entire research iceberg stretching from Functional
  Reactive Programming to internet routing and distributed algorithms.

  I did manage to surmount this through kludging and compromise. But I do
  not know whether these problems are merely a consequence of some design
  decision I could change to escape it, or if they are intrinsic to my
  (modest) UI goal.

  \begin{quote}
  This section reads more as an implementation detail - even though it
  makes some interesting points about the physics \& rods system. However,
  it might better to have it in a section that's more about implementation
  than about general principles. Or perhaps you could restructure Section
  3 to have (i) example (ii) general lessons (like physics etc.) and (iii)
  specific points about implementation (e.g.~your concrete physics system
  with rods, what you had to `polyfill')
  \end{quote}

  \hypertarget{visible-coordinate-systems}{%
  \subsubsection{Visible Coordinate
  Systems}\label{visible-coordinate-systems}}

  Rigidity in a flat world of sibling shapes is somewhat straightforward.
  However, rigidity in the SVG world is more involved.

  First of all, SVG shapes (e.g.~\texttt{\textless{}rect\textgreater{}})
  are strictly leaf nodes, so if I wish to nest boxes within boxes, the
  visible ``box'' must be a mere \emph{accessory} to the nestable element,
  in my case a \texttt{\textless{}g\textgreater{}} (group). This means
  that instead of resizing the x, y, width, height attributes of the
  \texttt{\textless{}rect\textgreater{}}, only its width and height change
  along with the transform attribute of its \emph{parent}
  \texttt{\textless{}g\textgreater{}}. This was not too bad; just
  subscribe this attribute, instead of the
  \texttt{\textless{}rect\textgreater{}} position, to the top-left Point
  handle.

  All child elements of a node transform with it, so already SVG has baked
  in a basic facility for translational rigidity. This is only available
  as a tree hierarchy, but it is still useful. However, it conflicts with
  my early decision to have Point objects all share the global co-ordinate
  system. This way, connecting ``this point'' to the mouse pointer, etc.
  is not infuriating to express. Still, it was necessary in the case of
  certain elements --- especially those which must transcend the tree
  structure altogether, like arrows between boxes --- to bite this bullet,
  one way or another.

  \begin{quote}
  Maybe mention that this is caused by the mismatch between the substrate
  and the programming system (substrate is tree, but programming is
  graphs)?
  \end{quote}

  Again, I return to how we tend to work things out in the freedom of
  paper. Co-ordinate systems, here merely positionally displaced, have
  their origins here and there and have vectors between them. The rods
  thus far let me visually express relations between global Points; now
  was a question of expressing one global Point as a displacement from
  another (the \texttt{\textless{}g\textgreater{}} transform). New rod
  Observables \texttt{p2\_from\_p1} and \texttt{p1\_from\_p2} do the
  vector subtraction, which can then be propagated as local co-ordinates
  to children. It is nicer to express the relation (as well as see it!)
  this way --- see Figure ?? above.

  \hypertarget{context-appropriate-ontologies}{%
  \subsubsection{Context-appropriate
  ontologies}\label{context-appropriate-ontologies}}

  Each API has its own conventions, including a way of naming and
  structuring expressions --- an ontology. The one-size-fits-all approach
  is exemplified in such interfaces. For example, in SVG we express a
  rectangle by
  \texttt{\textless{}rect\ x="10"\ y="10"\ width="600"\ height="400"\textgreater{}}.
  The SVG specification \emph{defines} to the user that a rect \emph{is} a
  top-left corner, a width, and a height --- and that's it. However, this
  SVG-approved parametrisation of a rectangle is far from the only
  one(1){[}https://www.shift-society.org/salon/papers/2018/critiques/critique-semprola.pdf{]},
  and thus is, unsurprisingly, ill-fitted to some contexts.

  For example, I find it natural to resize boxes by dragging any of their
  four corners, so I wanted this in OROM/SVG. In this context, a
  ``rectangle'' is \emph{seen as} four points: top-left, top-right,
  bottom-left, bottom-right. Obviously this is not a \emph{minimal}
  description, since given e.g.~the top-left and bottom-right, the other
  two points can be inferred. But one way or another, to be able to drag
  any of them, all four points must be present at \emph{some} level. This
  alternative ontology was polyfilled in the form of a ``rect controls''
  class that can be attached manually to any SVG
  \texttt{\textless{}rect\textgreater{}}. The SVG \texttt{x} and
  \texttt{y} are subscribed to the top-left Point; \texttt{width} and
  \texttt{height} are subscribed to Rod lengths.

  Another example is to be found in the DOM's event listener model.
  Conceptually, many ``events'' are in fact changes to the state of some
  physical (or conceptual) device. Some of these genuinely are different.
  But both keyboard keys and mouse buttons, for example, have two states
  --- pressed, not pressed --- which ought to make them interchangeable to
  an extent. This is essentially the logic behind ``key remapping'' in
  many PC games. But the official ontology of the Web makes it nontrivial
  to do this.

  To start with, the situation is modelled not as a time-varying value
  changing, but instead as a sort of Cartesian product of event listeners.
  Rather than, say, a \texttt{left\_mouse\_button} piece of live-state, we
  get \texttt{onmousedown} and \texttt{onmouseup}. Thankfully we do not
  have \texttt{onAdown}/\texttt{onAup}, \texttt{onBdown/onBup}, \ldots{}
  all the way to \texttt{onZdown}/\texttt{onZup}, just
  \texttt{onkeydown}/\texttt{onkeyup}. But this is just one of the many
  possible ways to slice this 3-dimensional\footnote{Device (mouse,
    keyboard), sub-device (button, key), state (up, down)} space.

  In OROM/SVG I did not quite want to re-map keys, but I did often want to
  have things follow the mouse when dragged. In order to do this, I
  reified the mouse pointer and its position, letting me write
  \texttt{subscribe(point.position,\ pointer.position)}. The Point has an
  \texttt{is-considering-me?} Observable wrapping
  \texttt{onmouseover}/\texttt{onmouseout}, and the LMB\footnote{left
    mouse button} is reified into \texttt{left\_mouse\_button\_is\_down}
  to be explicit. The aforementioned subscription is set up whenever
  \texttt{is-considering-me?} and \texttt{left\_mouse\_button\_is\_down}
  become true, and torn down otherwise. I tended to think of this in the
  form ``subscribe to pointer \emph{only when} pointer is-considering-me
  \emph{and} LMB is down'', but I could live with this as a future
  polyfill in JavaScript.

  It takes some frustration and experience to get used to the idea that
  you have a right to polyfill in an alternative representation. Before I
  came to this conclusion, I used to twist my head around translating my
  intention into the \texttt{x,y,width,height} parameters, and
  un-translating when reading back the code I had produced. Once you get
  used to having to adapt your mental imagery to a single way of doing
  things (i.e.~learning to code), it makes sense to simply expect to see
  more of it --- especially when you know you are new, surrounded by
  veterans who see no problem, etc. But nowadays I take the position that
  these are simply \emph{widespread} failings of our way of doing things,
  instead of \emph{our} failure to adapt to the way software is.

  What would it look like to support multiple ontologies? There are the
  two answers: anticipate the possibilities ahead of time, or support
  users adding their own. The latter seems to imply polyfilling. As for
  the former: in general, anticipating the diversity of ways someone might
  look at the world is doomed to fail. But in the case of fairly
  \emph{formalised} concepts such as geometrical shapes or mathematics, I
  could suggest that simple \emph{under-specification} is the root of the
  problem in SVG. Instead of explaining in English that ``\texttt{x} and
  \texttt{y} are the co-ordinates of the top left-hand corner\ldots{}'',
  it might be better to make these relations machine-readable or
  \emph{embodied} in the API. For the rect this might look like:

  A \texttt{rect} is\ldots{} - a \texttt{polygon} (defined elsewhere) -
  defined by 4 \emph{degrees of freedom} \texttt{x\ y\ width\ height}
  (internal representation) - where there are 4 \emph{vertices}, all
  \texttt{point}s + \texttt{{[}x,y{]}} ``top-left'' +
  \texttt{{[}x+width,y{]}} ``top-right'' + \texttt{{[}x,y+height{]}}
  ``bot-left'' + \texttt{{[}x+width,y+height{]}} ``bot-right''

  The hope is that if we then specify enough information --- say, the
  bot-left and top-right --- then the runtime has enough information to
  derive its internal 4 degrees of freedom. In a crude sense I have
  successfully anticipated the most obvious ontologies of a rectangle
  here. However, I missed out the ``centre plus half-width and
  half-height'' formulation, among many others.

  Is this a futile effort even for precise mathematical knowledge
  structures, or could enough formalisation of Euclidean geometry in the
  Web platform put an end to this sort of polyfilling?\footnote{This
    relates to my belief that type systems and other auto-reasoning tools
    are trapped in a doomed quest for ``closed-form'' AI, representable as
    a LaTeX formula.}

  \hypertarget{extensional-functions}{%
  \subsubsection{Extensional Functions}\label{extensional-functions}}

  Time and time again we come across the same pattern of partitioning
  system state: trees or graphs of dictonaries, a.k.a. Maps, a.k.a.
  associative arrays. I am talking about filesystem paths
  \texttt{/path/to/some/file}, Python / Java modules
  \texttt{com.example.pkg.subpkg}, JavaScript objects
  \texttt{window.my\_obj.component}. The common case is an association of
  a textual (string) name, to an arbitrary value. I find it useful to see
  this as a mathematical ``function'' defined \emph{extensionally} by
  listing its input/output mappings -- this opposed to an
  \emph{intensional} definition such as
  \texttt{x\ \textbackslash{}mapsto\ 2x+3}, or a computer program.

  Extensional functions are perhaps the most basic form of Knowledge
  Representation, and match natural language very well. ``The bicycle's
  wheels' spokes are silver'' straightforwardly translates to a function
  equation
  \texttt{root\ (bicycle)\ (wheels)\ (spokes)\ (colour)\ =\ root\ (silver)}.
  That is, whatever object is the output of \texttt{silver} in the
  top-level \texttt{root} function, the output of \texttt{colour} (in the
  function on the left) points to the same object. The ability to
  partition a system in this way enables what
  (2){[}https://github.com/amb26/papers/blob/master/ppig-2016a/ppig2016a.pdf{]}
  calls a ``natural co-ordinate system'' for a piece of software, crucial
  for understanding and adaptability by others (so long as it is
  externalised).

  It seems that this way of expressing the ``parts'' of a system is an
  inevitable requirement of any programming substrate. Some languages,
  such as C, do have static, \emph{compile-time} associative arrays
  (\texttt{struct}s). In my experience this is not enough, and it's
  necessary to bring in a library or clutter the code with a home-grown
  approximation. Some parts of the OROM authors' C code were confusing
  until I realised they were just the guts of a basic associative-array
  implementation; when I switched to JavaScript, these lines vanished.
  Perhaps another strength of JavaScript is its low concrete syntax cost
  for \emph{instances} or \emph{literals} of associative arrays. Writing
  or reading

  \begin{lstlisting}[language=JavaScript]
  a = {
    b1: { c1: z, c2: y },
    b2: { c3: x, c4: w }
  };
  \end{lstlisting}

  is more WYSIWYG\footnote{What You See Is What You Get} than the
  imperative-style

  \begin{lstlisting}[language=JavaScript]
  a = new Map();
  a.set('b1', new Map());
  a.get('b1').set('c1', z);
  a.get('b1').set('c2', y);
  a.set('b2', new Map());
  a.get('b2').set('c3', x);
  a.get('b2').set('c4', w);
  \end{lstlisting}

  The latter is unfortunately still required in JavaScript for extensional
  functions with \emph{non-string} inputs.

  The decision to switch from the flat tabular representation in OROM/HTML
  to the nestable, Boxer-like structure in OROM/SVG permitted extensional
  (tree) functions in its substrate. True, I could always encode these in
  flat tables by having them reference each other. But this is like the
  second listing above: hard to follow. In the case of distinguishing
  between ordinary state mappings and the ``method dictionary'' of
  vtables, I actually stored method mappings with a ``-'' character to
  avoid doing this.

  (Screenshot showing this)

  In OROM/SVG, I can directly express the model I was thinking of: the
  vtable box has a \texttt{methods} box and that's where the method boxes
  go.

  (Screenshot showing this)

  \hypertarget{externalisability}{%
  \subsubsection{Externalisability}\label{externalisability}}

  This refers to exposing structured program state to the user. This data
  can then be saved and used to restore the system at a later date, but it
  can also be tweaked with corresponding changes reflected in the system.

  Externalisability, or its weaker cousin \emph{persistence}, was
  absolutely necessary to continue OROM/SVG development past a certain
  point.

  Reference Antranig's stuff on this. How I dealt with event listeners etc

  Necessity to development: frustrating to find bug, tear down, refresh,
  lose work, build it up again!

  Inspect Element: copy to save works, arbitrary edits \emph{should} work
  (cuz it's on demand)

  Work required to parse SVG trees into objects ``on demand''

  Started heavyweight, ended lightweight (generate rods on demand, let SVG
  move children)

  \hypertarget{polyfilling}{%
  \subsubsection{Polyfilling}\label{polyfilling}}

  Very advantageous to permit this e.g.~JavaScript

  Stupid lack of XYZ in JS or SVG due to come out in the next spec, etc

  Syntax sugar (SVG tree ops, vector ops)

  objects in JS user data

  \hypertarget{the-orom-system-as-a-part-of-the-solution}{%
  \section{The OROM system as a part of the
  solution}\label{the-orom-system-as-a-part-of-the-solution}}

  Meta-circularity as empowerment, not just ``cool''

  The allure

  Other themes e.g.~homoiconicity of code and data

  Problems: cognitive vs.~``official'' complexity

  Maxwell's eqns metaphor: omits the years of study compressed into the
  math symbols!

  And yet, the allure of the Maxwell's Eqns of Software?

  ``Taming'' SVG - ability to use SVG in a vaguely sane way (vector
  graphics editor)

  Similarly for e.g.~3D OpenGL, Web Audio, sockets \ldots{}

  \hypertarget{tensions-between-philosophies}{%
  \subsection{Tensions between
  philosophies}\label{tensions-between-philosophies}}

  \hypertarget{stable-or-emergent-antranig-against-oop-dispatch}{%
  \subsubsection{Stable or emergent? (Antranig against OOP
  dispatch)}\label{stable-or-emergent-antranig-against-oop-dispatch}}

  \hypertarget{turing-complete-antranig-against-this}{%
  \subsubsection{Turing-complete? (Antranig against
  this)}\label{turing-complete-antranig-against-this}}

  \hypertarget{messaging-vs-read-write}{%
  \subsubsection{Messaging vs Read-Write?}\label{messaging-vs-read-write}}

  \hypertarget{misc}{%
  \section{Misc}\label{misc}}

  A note about DIY and libraries

  My other problem was one of cost-benefit. My aim was to \ldots{}

\acks
I want to thank ....

\printbibliography
\end{document}