### 2020-01-21 convivial-2020/tex/manuscript.tex

\documentclass[english,submission]{programming}
\usepackage[backend=biber]{biblatex} % Use Biblatex
\addbibresource{example.bib}

\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete, do, else,
  finally, for, function, if, in, instanceof, new, return, switch, this, throw,
  try, typeof, var, void, while, with, let, false, true},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  sensitive=true
}
\lstset{language=JavaScript, numbers=none}

\providecommand{\tightlist}{}% Don't want Pandoc's tight lists
  %\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\paperdetails{
  perspective=art,
  area={General-purpose programming}
}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002944.10011122.10003459</concept_id>
<concept_desc>General and reference~Computing standards, RFCs and guidelines</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{General and reference~Computing standards, RFCs and guidelines}`

\input{../manuscript.tex}

\acks
I want to thank ....

\printbibliography
\end{document}