### 2020-01-22 convivial-2020/tex/manuscript.tex

\documentclass[english,submission]{programming}
\usepackage[
backend=biber,
style=alphabetic
]{biblatex} % Use Biblatex
\addbibresource{../refs.bib}

\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete, do, else,
  finally, for, function, if, in, instanceof, new, return, switch, this, throw,
  try, typeof, var, void, while, with, let, false, true},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  sensitive=true
}
\lstset{language=JavaScript, numbers=none}

\providecommand{\tightlist}{}% Don't want Pandoc's tight lists
  %\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\paperdetails{
  perspective=art,
  area={Domain-specific programming, visual programming, self-sustaining systems}
}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011050.10011017</concept_id>
       <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011050.10011058</concept_id>
       <concept_desc>Software and its engineering~Visual languages</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Domain specific languages}
\ccsdesc[300]{Software and its engineering~Visual languages}

\input{../manuscript.tex}

\printbibliography
\end{document}