### 2020-05-15 convivial-2020/acm-style/ccs20-acm.tex
All day --> ACM style. I hate BibTeX. heck maybe i just hate latex altogether

\documentclass[sigconf,screen]{acmart}
\usepackage{listings}

\citestyle{acmauthoryear}

\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete, do, else,
  finally, for, function, if, in, instanceof, new, return, switch, this, throw,
  try, typeof, var, void, while, with, let, false, true},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  sensitive=true
}
\lstset{language=JavaScript, numbers=none}

\providecommand{\tightlist}{}% Don't want Pandoc's tight lists
  %\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\setcopyright{acmlicensed}
\acmPrice{15.00}
\acmDOI{10.1145/3397537.3397549}
\acmYear{2020}
\copyrightyear{2020}
\acmSubmissionID{prog20ccs-id8-p}
\acmISBN{978-1-4503-7507-8/20/03}
\acmConference[<Programming'20> Companion]{Companion Proceedings of the 4th International Conference on the Art, Science, and Engineering of Programming}{March 23--26, 2020}{Porto, Portugal}
\acmBooktitle{Companion Proceedings of the 4th International Conference on the Art, Science, and Engineering of Programming (<Programming'20> Companion), March 23--26, 2020, Porto, Portugal}

\begin{document}
% The following ONLY works after the \begin{document}!!
\renewcommand{\textasciitilde}{~}% Don't want Pandoc's BS with the ~ space

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011050.10011017</concept_id>
       <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011050.10011058</concept_id>
       <concept_desc>Software and its engineering~Visual languages</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Domain specific languages}
\ccsdesc[300]{Software and its engineering~Visual languages}

\input{manuscript-acm.tex}

\bibliography{../refs.bib}

\end{document}