### 2021-06-13 plop-2021/plop21-master.tex
archive and tidyup work so far

\documentclass{article}
\usepackage{epsfig}
\usepackage{booktabs}
\usepackage{ucs}
\usepackage[utf8x]{inputenc}

% The titlebox defines how much vertical space is given for
% the authors' list. If you need extra space to show all the
% authors, uncomment the line below and increase the value. Please
% do not make the titlebox smaller than the original size of 5cm.
%\setlength\titlebox{5cm}

\title{Technical Dimensions of Programming Systems}

\author{Joel Jakubovic}
\affiliation{%
  \institution{University of Kent}
  \city{Canterbury}
  \country{UK}}
\email{jdj9@kent.ac.uk}
\author{Jonathan Edwards}
\affiliation{}
\email{jonathanmedwards@gmail.com}
\author{Tomas Petricek}
\affiliation{%
  \institution{University of Kent}
  \city{Canterbury}
  \country{UK}}
\email{T.Petricek@kent.ac.uk}

\authorsaddresses{}

\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

\newcommand{\joel}[1]{}
\newcommand{\note}[1]{}
\newcommand{\tp}[1]{}
\newcommand{\hypertarget}[1]{}

\input{plop21-manuscript.tex}

\bibliography{plop21}
\bibliographystyle{apacite}
\end{document}