### 2021-06-15 plop-2021/plop21-master.tex
tidy for initial submission

\documentclass[acmlarge, review=true]{acmart}

\acmConference{PLoP}{2021}{online}

\markboth{}{}

\title{Technical Dimensions of Programming Systems}

\author{Joel Jakubovic}
\affiliation{
\institution{University of Kent} \city{Canterbury} \country{UK}
}
\email{jdj9@kent.ac.uk}
\author{Jonathan Edwards}
\affiliation{}
\email{jonathanmedwards@gmail.com}
\author{Tomas Petricek}
\affiliation{
\institution{University of Kent} \city{Canterbury} \country{UK}
}
\email{T.Petricek@kent.ac.uk}

\begin{abstract}
Programming systems often go beyond programming languages. Programming
is often done in the context of a stateful environment, through a
graphical user interface, by interacting with the system rather than
just by writing code. Much research effort focuses on building
programming systems that are easier to use, accessible to non-experts,
moldable and/or powerful, but such efforts are often disconnected. They
are informal, guided by the personal vision of the authors and thus are
only evaluable and comparable on the basis of individual experience
using them. In other words, they fail to form a coherent body of
research. It isn't clear how to build on what has been done. Much has
been said and done that allows comparison of \emph{programming
languages}, yet no similar theories exist for \emph{programming
systems}. We believe that programming systems deserve a theory too. We
examine some influential past programming systems and review their
stated design principles, technical capabilities and styles of user
interaction. We propose a framework of \emph{technical dimensions},
which captures the underlying system characteristics and provide a means
for conceptualising and comparing programming systems. We thus break the
holistic view of programming systems down into characteristics that can
be compared or advanced independently. It makes it possible to talk
about programming systems in a way that can be shared and agreed or
constructively disagreed upon, rather than relying just on individual
experience of using them. The aim is to bring programming systems into
the focus of programming research by providing the foundations that
allow such research to proceed in a more systematic way. In other words,
we want to allow the designers of future programming systems to
(finally) stand on the shoulders of the giants.
\end{abstract}

\begin{document}

\maketitle
%\thispagestyle{empty}

\newcommand{\joel}[1]{}
\newcommand{\note}[1]{}
\newcommand{\tp}[1]{}
\providecommand{\tightlist}{}% Don't want Pandoc's tight lists
%\newcommand{\hypertarget}[1]{}

\input{plop21-manuscript.tex}

\appendix

\input{plop21-appendix.tex}

\citestyle{acmauthoryear}
\bibliographystyle{ACM-Reference-Format}
\bibliography{plop21}
\end{document}