### 2022-09-14 onward-2022/onward22-master.tex
Onward! '22 submission

\documentclass[sigplan,review]{acmart}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{amsthm}

\acmConference{Onward! Essays}{December 5--10, 2022}{Auckland, New Zealand}

\title{Ascending the Ladder to Self-Sustainability}
\subtitle{Achieving open evolution in an interactive graphical system}
\author{Joel Jakubovic}
\affiliation{\institution{University of Kent} \city{Canterbury} \country{United Kingdom}}
\email{jdj9@kent.ac.uk}
\author{Tomas Petricek}
\affiliation{\institution{Charles University} \city{Prague} \country{Czechia}}
\email{tomas@tomasp.net}

\begin{document}

\lstset{breaklines=true, columns=fullflexible, breakatwhitespace=true, basicstyle=\ttfamily\footnotesize}
\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete, do, else, finally, for, function, if, in, instanceof, new, return, switch, this, throw, try, typeof, var, void, while, with, let, const},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  sensitive=true
}

\newcommand{\joel}[1]{}
\newcommand{\tomas}[1]{}
\newcommand{\note}[1]{}
\newcommand{\notes}[1]{}
\newcommand{\todo}[1]{\textbf{TODO: #1}}
\newcommand{\delete}[1]{\textbf{DELETE:}#1}
\providecommand{\tightlist}{}% Don't want Pandoc's tight lists
% Nor do we want long tables by default
\newenvironment{longtable}[2]{\begin{tabular}}{\end{tabular}}
\newenvironment{head}{}{}
\theoremstyle{definition}
\newtheorem{force}{Force}
\newtheorem{heuristic}{Heuristic}
\newtheorem{requirement}{Requirement}
\newcommand{\naive}{na\"ive}
%\newcommand{\hypertarget}[1]{}

\input{onward22-manuscript.tex}

\appendix
\input{onward22-appendix.tex}

\bibliographystyle{plainurl}
\bibliography{onward22}
\end{document}