### 2022-09-27 prog-2022/changelog.tex
New cover letter, ensure nbsp pre cites

\section*{Technical Dimensions of Programming Systems: major revisions}

We thank the reviewers for their continued attention and feedback. The
major changes we have made are as follows.

\paragraph{Definition of ``system''.}

We worked out a sensible explicit definition of what a programming
system is and included it as early as possible in the paper.

\paragraph{L/O/A typology.}

We intended the three-way classification in section~3 merely to provide
some rhyme and reason as to where our example systems were coming from.
We did not intend it as a rigid scheme where membership in a category
means we are making some profound or bold claim. The L/O/A-type naming
seems to have misled readers into anticipating the latter, so we have
relaxed the naming to ``language-based'', ``OS-like'' and
``application-focused''. We did refer back to this scheme in one
specific context, in section~5.2, noting that most OS-likes (Unix, Lisp
and Smalltalk) cluster in a region to the right.

\paragraph{Dark \& evaluation.}

Section~5 has been clarified as an evaluation of the framework by means
of the two example usages. We've included a summary table for the
dimensional analysis of Dark in section~5.1 and annotated the screenshot
for convenience. We appreciate the reviewer's desire for examples of the
specific interactions, but we thought it'd be pointless to write down
something already explained in one of Dark's tutorial videos, so we
cited the latter instead.

\paragraph{References to strengthen claims.}

Our statement about ``cool/interesting'' was entirely from experience,
so we relaxed it to something we could more easily support with
references. These can be seen in a footnote with our remarks linking
them to the point we're trying to make. Elsewhere we have addressed this
by deleting the claim (e.g.~for ``holistic view'') or adding references,
for example to the various example systems like Java and Jupyter. We're
afraid we don't understand the request to put explicit examples in the
section on venues for publishing programming systems research; we'd
already listed UIST, VL/HCC, LIVE and PX. If this was about specific
publications instead, we've now referred to some in the footnote on page
3.

\paragraph{Socio-Technical dimensions?}

The fact of the matter is that almost all of our 22 dimensions are
technical, with the possible exception of Learnability and Sociability
being socio-technical, and some remarks in Integrity vs.~Openness. On
this basis, we strongly disagree with calling the contribution
``socio-technical'' dimensions or similar, as this would give a false
expectation to readers. We appeal to the reviewer to permit us to keep
the name ``technical''---we put a lot of effort into thinking of names
early on, but ``technical dimensions'' was the best out of perhaps only
two viable choices. If this is still unacceptable, we will fall back on
the other name, ``design dimensions'', but if this is still rejected we
would need some help determining an alternative.

\paragraph{On the new length of the paper.}

Following approval from the editors, this revision includes the full
catalogue of dimensions in the main body of the paper. While we
acknowledge this makes it very long, we do not intend reviewers or
readers to read this section from start to finish. Instead, it is meant
as a reference to be used as needed, for example to get further detail
on the dimensions we reference later for the Dark programming system.
This lets us present the material as we originally intended and
addresses a reviewer concern that it was awkward to refer to appendix
material in the section on Dark. In summary, we always saw the
dimensions catalogue as belonging in the main body, but for review
purposes it may still be treated like an appendix.